---
title: "Homework 1"
author: "Jose Saavedra"
date: "2/22/2018"
output: github_document
---



##Problem 1

The data dictionary can be found in the hw01 folder under imports-85-dictionary.txt


##Problem 2

Importing Data
```{r problem 2}
names_columns <-  c('symboling',
                    'normalized_losses'
                    ,'make','fuel_type'
                    ,'aspiration',
                    'num_of_doors',
                    'body_style',
                    'drive_wheels',
                    'engine_location',
                    'wheel_base',
                    'length',
                    'width',
                    'height',
                    'curb_weight',
                    'engine_type',
                    'num-of_cylinders',
                    'engine_size',
                    'fuel_system',
                    'bore',
                    'stroke',
                    'compression_ratio',
                    'horsepower',
                    'peak_rpm',
                    'city_mpg',
                    'highway_mpg',
                    'price'
                    )
columns_types <- c('integer',
                   'real',
                   'character',
                   'character',
                   'character',
                   'character',
                   'character',
                   'character',
                   'character',
                   'real',
                   'real',
                   'real',
                   'real',
                   'integer',
                   'character',
                   'character',
                   'integer',
                   'character',
                   'real',
                   'real',
                   'real',
                   'integer',
                   'integer',
                   'integer',
                   'integer',
                   'integer'
                   )
autos <- read.csv(file = 'imports-85.data',
                  col.names = names_columns,
                  colClasses = columns_types,
                  na.strings = '?',
                  header = F)
str(autos)

library(readr)


autos2 <- read_csv(file = 'imports-85.data',
                   col_names = names_columns,
                   na='?',
                   col_types = list(
                   normalized_losses=col_double()
                   ))
str(autos2)
```

##Problem 3

Technical Questions about importing data

a.) The first row becomes the column names, this is bad you can fix this by adding header=F.

b.)The code gives use all the data with generic column names V# 

c.)they become factors since the columns have at least one character, and a column of characters in dataframe are automatilly made as factors 

d.)The main reason the second option is larger is due to the fact taht the first frame saves the characters as factors, which are stored in a ledger and called on when seen, where as in the second frame we store the charaters as unique characters so they take up more space.

e.)All character types are coerced as NA, and the integers are converted to reals

##Problem 4

Practice base plotting


```{r  Problem 4}
hist(autos$price, col = c('blue','red', 'green'),xlab = 'Price in Dollar',
     main = 'Histogram of Prices of Autos')

boxplot(autos$horsepower, horizontal = T, main='Box plot og horsepower')

barplot(sort(table(autos$body_style), decreasing = T), col = c('red','blue'))

stars(autos[autos$aspiration=='turbo',][,c(10:13,26)],main = 'stars plot of cars with turbo',key.loc = c(18,3))
```


##Problem 5

Summaries

a.)
```{r problem 5a}
mean(autos[autos$fuel_type =='gas',]$price, na.rm = TRUE)

mean(autos[autos$fuel_type=='diesel',]$price, na.rm = TRUE)
```

b.)
```{r problem 5b}
autos[autos$num.of_cylinders=='twelve',]$make
```

c.)
```{r problem 5c}
A <- aggregate(autos[,"make"], by=list(autos$make,autos$fuel_type),FUN = length)
A[which.max(A[A$Group.2=='diesel',]$x),]$Group.1

```

d.)
```{r problem 5d}
autos[which.max(autos$horsepower),]$price
```
Note that the car with the largest horsepower has an unknown price and therefore we get the above result (car 130, make porsche)

e.)
```{r problem 5e}
Bottom10 <- range(autos$city_mpg, na.rm = T)
Bottom10[2] <- quantile(autos$city_mpg, probs = .1)
Bottom10
```
the output in the range where the bottom 10% lie

f.)
```{r problem 5f}
Top10<- range(autos$highway_mpg, na.rm = T)
Top10[1] <- quantile(autos$highway_mpg, probs = .9)
Top10
```
the output is the range where the top 10% lie

g.)
```{r problem 5g}
median(autos[autos$city_mpg<=Bottom10[2],]$price,na.rm = T)
```

##Problem 6

Technical Questions about data frames

a.) `dat$column_name` allows us to call on the column of name: column_name, and treating this as a vector `dat$xyz` if xyz does not exist then R returns NULL

b.) Number 4 fails 

c.)It fails since R doesn't understand that mpg is a name, this has to be explicitly denoted with a $ or "" or use row, column notation []

d.)Yes since a data frame is at its basic structure a list, and you can append a list with a list since they are not atomic structures.

e.) returns a list of 11 numerical vectors with the associated names

f.) `as.data.frame()` can convert a list to a list into a data frame


#Problem 7

```{r problem 7}
library(corrplot)
qdat <- cor(na.omit(autos[,c(10:14, 17, 19:26)]))
#corrplot number 1
corrplot(qdat, method = 'circle')
#corrplot number 2
corrplot(qdat, method = 'color', type = 'upper')
```
Things that I noted, any descriptors to do with size are all postively correlated and anything to do with mileage and proformace is negatively correlated


##Problem 8

I went to office hours where the professor told discovered that he didnt mean to use qdat so the following is what we were supposed to do

###8.1

```{r Problem 8.1}
PCA_autos <- prcomp(na.omit(autos[,c(10:14, 17, 19:26)]), scale. = T)
eigen_values <- PCA_autos$sdev^2
barplot(eigen_values, names.arg = colnames(PCA_autos$x))
```
The first 3 Principal components capture 
```{r Problem 8.1b}
sum(eigen_values[1:3]/sum(eigen_values))*100
```

###8.2

a.)
```{r 8.2a}
PC_1 <- PCA_autos$x
plot(PC_1[,1],PC_1[,2],cex=1, xlab = 'Principal Component 1', ylab = 'Principal Component 2')
```
b.)
```{r 8.2b}
PC_loadings <- PCA_autos$rotation
plot(PC_loadings[,1],PC_loadings[,2],pch=19,cex=.7,
     xlab = "PC #1 Loading",ylab = "PC #2 Loading",
     xpd=T, xlim = c(-.5,.6))
text(PC_loadings[,1],PC_loadings[,2]+.03, labels = row.names(PC_loadings),cex = 1,xpd=T)
```

c.) 
```{r 8.2c}
biplot(PCA_autos,scale = 0)
```

d.) From the loadings we can see that mpg variables are closely related, while the variables dealing with size are close.




